Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-07-31 SDSS Journal Paper/Paper Draft.Rmd
adding proper citations within paper

# Introduction
As the amount of data has increased in nearly every facet of life, the need to make sense of that data in an approachable, accessible form has become ever more important. 
As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form \citep{fisher}.
<!-- With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \cite{fisher}.  -->
In many cases, dashboards support viewers' information processing, helping to make sense of complex data, navigate through a dataset, and supporting decision making based on the data.
<!-- A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making.  -->
Dashboards are often used, as with the car display of the same name, to provide summary information about many separate attributes of a common entity. One glance at a car's dashboard will tell you the speed, RPM, engine temperature, amount of gas in the tank; more importantly, however, the goal is not for the user to remember all of these characteristics, but to assess whether any of these quantities is outside of the expected range. 
Similarly, interactive dashboards for data are often used to display many different attributes and performance metrics which are of importance for stakeholders.
<!-- An important area of dashboard design and usage is to visualize many different attributes that are useful to display key performance metrics of importance for stakeholders.  -->
In this paper, we discuss the process of designing a dashboard to present publicly available government data to stakeholders in small Iowa towns to facilitate decision making and objective comparison with other similarly-situated towns.

@@

The Smart and Connected Community (SCC) dashboard data are primarily assembled from  \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources. Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level. In addition to the more commonly available statistics derived from e.g. the census and American Community Survey, \url{data.iowa.gov} contains several unique data sets, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.

Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.
Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school. The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education. \db{[@petersCommunityResiliencyDeclining2019] -- Change the reference}
<!-- \svp{Cite Dave's papers about the ISTP QOL metrics, including education}
 These data are used for details are a town level, that will provide information about the quality of life. -->

@@

Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary. As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest. This design is based on research into visualization sensemaking [@lee], in that we allow users to explore outward from the familiar to the unknown. \db{The map visuals were built using Open Source Routing Machine (OSRM) route functions [@luxen-vetter-2011] in R [@r] to amplify the accuracy of the distances from necessary services in town-centric point. OSRM allows for finding the "As the Crow Flies" distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.}

@@

\svp{Users' towns} \db{Town Users' ??} are at the center of our application, and comparisons to other, similar towns are secondary. As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest. 
<!-- The suggested comparison towns are generated using unsupervised clustering methods on different sets of variables, but as we acknowledge that the problem of what towns should be used for comparisons is a tough one even for humans, we still want to allow for user input in this space.  -->

>>>>

# Introduction
As the amount of data has increased in nearly every facet of life, the need to make sense of that data in an approachable, accessible form has become ever more important. 
As a result, many companies and organizations use interactive dashboards to present these data in a more useful and visually appealing form [@fisher].
<!-- With the amount of publicly open-source data, a proliferation of visualization dashboards has increased in nearly every industry \cite{fisher}.  -->
In many cases, dashboards support viewers' information processing, helping to make sense of complex data, navigate through a dataset, and supporting decision making based on the data.
<!-- A dashboard in its fundamental form, a dashboard supports a way of presenting and making sense of complex data to better enable and support decision making.  -->
Dashboards are often used, as with the car display of the same name, to provide summary information about many separate attributes of a common entity. One glance at a car's dashboard will tell you the speed, RPM, engine temperature, amount of gas in the tank; more importantly, however, the goal is not for the user to remember all of these characteristics, but to assess whether any of these quantities is outside of the expected range. 
Similarly, interactive dashboards for data are often used to display many different attributes and performance metrics which are of importance for stakeholders.
<!-- An important area of dashboard design and usage is to visualize many different attributes that are useful to display key performance metrics of importance for stakeholders.  -->
In this paper, we discuss the process of designing a dashboard to present publicly available government data to stakeholders in small Iowa towns to facilitate decision making and objective comparison with other similarly-situated towns.

@@

The Smart and Connected Community (SCC) dashboard data are primarily assembled from  \url{data.iowa.gov} [@iowa_gov], with some additional datasets assembled from federal and private sources. Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level. In addition to the more commonly available statistics derived from e.g. the census and American Community Survey, \url{data.iowa.gov} contains several unique data sets, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.

Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.
Data collected from ELSI [@elsi] from \url{https://nces.ed.gov} were used to show the distance to any private or public school. The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education [@zarecor2021rural].
<!-- \svp{Cite Dave's papers about the ISTP QOL metrics, including education}
 These data are used for details are a town level, that will provide information about the quality of life. -->

@@

Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary. As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest. This design is based on research into visualization sensemaking [@lee], in that we allow users to explore outward from the familiar to the unknown. The map visuals were built using Open Source Routing Machine (OSRM) route functions [@luxen-vetter-2011] in R [@r] to amplify the accuracy of the distances from necessary services in town-centric point. OSRM allows for finding the "As the Crow Flies" distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.

@@

Users' towns are at the center of our application, and comparisons to other, similar towns are secondary. As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest. 
<!-- The suggested comparison towns are generated using unsupervised clustering methods on different sets of variables, but as we acknowledge that the problem of what towns should be used for comparisons is a tough one even for humans, we still want to allow for user input in this space.  -->

### 2022-07-31 SDSS Journal Paper/Paper-Draft.tex
adding proper citations within paper

The Smart and Connected Community (SCC) dashboard data are primarily assembled from \url{data.iowa.gov}, with some additional datasets assembled from federal and private sources. Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level. In addition to the more commonly available statistics derived from e.g.~the census and American Community Survey, \url{data.iowa.gov} contains several unique data sets, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.

Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.
Data collected from ELSI from \url{https://nces.ed.gov} were used to show the distance to any private or public school. The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education. {\textcolor{blue}{[@petersCommunityResiliencyDeclining2019] -- Change the reference}}

@@

Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary. As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest. This design is based on research into visualization sensemaking \citep{lee}, in that we allow users to explore outward from the familiar to the unknown. {\textcolor{blue}{The map visuals were built using Open Source Routing Machine (OSRM) route functions [@luxen-vetter-2011] in R [@r] to amplify the accuracy of the distances from necessary services in town-centric point. OSRM allows for finding the "As the Crow Flies" distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.}}

@@

{\textcolor{red}{Users' towns}} {\textcolor{blue}{Town Users' ??}} are at the center of our application, and comparisons to other, similar towns are secondary. As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.

>>>>

The Smart and Connected Community (SCC) dashboard data are primarily assembled from \url{data.iowa.gov} \citep{iowa_gov}, with some additional datasets assembled from federal and private sources. Most of these data sets are collected at a town/city or county spatial resolution, requiring us to carefully join data to ensure that these differences are respected while collating relevant information at the city level. In addition to the more commonly available statistics derived from e.g.~the census and American Community Survey, \url{data.iowa.gov} contains several unique data sets, including local liquor sales, school building locations, town budgets and expenditures, hospital beds, Medicaid reimbursements, and other details that may provide information about local quality of life.

Data available on Iowa's data portal were augmented in some cases with higher-quality data sets in cases where the Iowa data were out of date or insufficiently accurate.
Data collected from ELSI \citep{elsi} from \url{https://nces.ed.gov} were used to show the distance to any private or public school. The National Center for Education Statistics (NCES) is the primary federal entity for collecting and analyzing data related to education \citep{zarecor2021rural}.

@@

Our primary objective is to provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary. As a result, the next component of the dashboard is intended to provide a brief overview of the information we have about a specific town of interest. This design is based on research into visualization sensemaking \citep{lee}, in that we allow users to explore outward from the familiar to the unknown. The map visuals were built using Open Source Routing Machine (OSRM) route functions \citep{luxen-vetter-2011} in R \citep{r} to amplify the accuracy of the distances from necessary services in town-centric point. OSRM allows for finding the ``As the Crow Flies'' distance and time on the road for our vital services map, since OSRM technology is similar to Google maps.

@@

Users' towns are at the center of our application, and comparisons to other, similar towns are secondary. As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.
